---
title: "Identifying fishing trips workflow"
author: "Tania Mendo"
date: "11/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Identifying individual fishing trips
The following is a conceptual workflow for identifying fishing trips from highly resolved spatial data. These data cab be obtained from different devices, from Automatic Identification Systems (AIS), to handheld GPS, or any trackers transmitting at least latitude, longitude, and time stamp.

Depending on the device, and the information available for each fishery, aspects of the workflow might change.

In the section below please just use your own data set and start with point 1!!!

```{r load dataset, echo = FALSE}

#to identify fishing trips:

#df<-read.table("test.txt",sep="\t",header = TRUE)
#df<-read.table("test.txt",sep=",",header = TRUE)
df<-read.csv("Tracks.csv")#314388
trips<-read.csv("Trips.csv")
devices<-read.csv("Devices.csv")
vessels<-read.csv("Vessels.csv")
df$trip_id<-as.factor(df$trip_id)
trips$trip_id<-as.factor(trips$trip_id)
#314388
df<-merge(df,trips[,c("trip_id","device_id")],all.x=TRUE)
df<-merge(df,devices[,c("device_id","vessel_id")],all.x=TRUE)
#303000
df$vessel_id<-as.factor(df$vessel_id)

df$time_stamp<-as.POSIXct(as.character(df$time_stamp), format = "%Y-%m-%d %H:%M:%S")

#how to add mili seconds?
#df$time_stamp2<-as.POSIXct(as.character(df$time_stamp), format = "%Y-%m-%d %H:%M:%OS3")
#cambiar resoucion a decimales de segundo

```

#1.-Define extent of study area

Remove latitudes and longitudes outside this extent. Rationale: Some devices might send 0,0 messages for lats and longs when there is not good reception of satellites. Also, if using on-board observers or fishers that transport the devices you might get points in land if they turned them on before the trip.

```{r extent}
library(tidyverse)

df<-df%>%
  filter(latitude>55)

df<-df%>%
  filter(longitude<0&longitude>(-8))

```

#2.- Remove duplicates

```{r duplicates, include=FALSE}

#Remove positions for each vessel that are recorded at the same time
df$vessel_id<-factor(df$vessel_id)

df2<-df %>% 
  distinct(vessel_id,time_stamp, .keep_all = TRUE)

#Remove observations per vessel occurring in the exact same location
df3<-df2 %>% 
  distinct(vessel_id,longitude,latitude, .keep_all = TRUE)
```

#3.- Remove points in land and X metres buffer

```{r load example dataset fishing trips, include=FALSE}

library(sf)
library(ggplot2)
library(raster)

df_sf = st_as_sf(df3, coords = c("longitude", "latitude"), crs = 4326)
df_sf_utm<-st_transform(df_sf, 32630)

coastline <- shapefile("OutHebrides.shp") 
coastline_sf <- st_as_sf(coastline)#in UTM
#coastline_sf_utm <- st_transform(coastline_sf, crs = 32630)
coastline_sf_10_utm<-coastline_sf%>%
  st_buffer(dist=10)#set a 10 metre buffer around land

points_in_buffer2<- df_sf_utm%>% rowid_to_column("seq") %>% st_intersection(coastline_sf_10_utm) %>% pull(seq)

df4<-df3 %>% 
  rowid_to_column("seq") 

df4 <- df4[which(!df4$seq %in% points_in_buffer2),]#this to tidyverse?

```


#**************************
#4a.- Construct trajectories
#**************************

```{r trajectories, include=FALSE}

library(adehabitatLT)
df4$date<-as.Date(df4$time_stamp)

df4$id<-paste(df4$vessel_id,df4$date)#here I am acknowledging that a trip will not be conducted overnight.
df4_sf<-st_as_sf(df4, coords = c("longitude", "latitude"), crs = 4326)

df4_sf_utm<-st_transform(df4_sf, 32630)
coords_utm<-as.data.frame(st_coordinates(df4_sf_utm))
df4$x<-coords_utm$X
df4$y<-coords_utm$Y

traj<-as.ltraj(df4[, c("x", "y")], df4$time_stamp,df4$id)#

trajdf<-ld(traj)#back to data frame

#**************************
#4b.- Joefine polygon's approach - define trips when entering and leaving a polygon?
#**************************

###
```

#5.- Remove points with unrealistic speeds

```{r speeds, include=FALSE}

#We know that in our fishery, speeds greater than 25 knots are not likely

trajdf$speed<-trajdf$dist/trajdf$dt*1.94384449#in knots

repeat { 
subset<-trajdf[trajdf$speed>25,]
subset<- subset[complete.cases(subset), ]
 
#sel<-as.data.frame(sel)
sel<-factor(subset$pkey)

nrows<-length(sel)

  if (nrows==0) {
    break
    } 
  else
trajdf<-trajdf[!trajdf$pkey %in% sel,] 
traj<-as.ltraj(trajdf[, c("x", "y")], trajdf$date,trajdf$burst)
trajdf<-ld(traj)
trajdf$speed<-trajdf$dist/trajdf$dt*1.943

} 
```

#6.- Remove "bad" trips

```{r trips, include=FALSE}

trips_summary<-trajdf%>%
  group_by(burst) %>%
  summarise(dist_traveled = sum(dist,na.rm=TRUE),time_travelled=sum(dt/3600,na.rm=TRUE), nobs=n_distinct(date))  

ggplot()+geom_histogram(data=trips_summary,aes(x=dist_traveled))+theme_bw()#in meters

ggplot()+geom_histogram(data=trips_summary,aes(x=time_travelled))+theme_bw()#in hours

sel<-trips_summary[trips_summary$time_travelled<1,]
sel<-factor(sel$burst)

trajdf2<-trajdf[!trajdf$burst %in% sel,] 

```

