---
title: "WKSSFGEO"
output: html_document
editor_options: 
  chunk_output_type: console
---

#Identify individual fishing trips

My first thoughts on this is that if people want to identify fishing trips then they will have information on fishing vessels but not about an actual trip identifier. at least this is the situation I deal with. Then, depending on the fishery, you might have fleets where it is really easy to identify trips, say they only go once a day, or it might be more complicated and you have to come up with rules to "define" what a fishign trip is. you might want to have a time between consecutive observations that can be used to "cut" daily data into two trips. 

I would suggest we start with this small number of columns and work our way with the same code, instead of having people maybe do different things beforehand which can add more complications. 

For identification of trips, the idea I think would be to apply rules to the dataset and then review if the trips identified (and also the ones ruled out) actually were the right ones. This can be done via visualization of trips or with on-board observer information?

```{r load example dataset fishing trips, include=FALSE}

#to identify fishing trips:

df<-read.table("example_dataset.txt",sep=",")
df$vessel_id<-as.factor(df$vessel_id)
df$time_stamp<-as.POSIXct(as.character(df$time_stamp), format = "%Y-%m-%d %H:%M:%S")

```

#Inferring fishing activities (hauling)

Here we want to infer when fishing activities (hauling events) are occurring. For this I think it is better if participants have some validated data to run different models and test which one performs better with their data.


```{r load example dataset fishing trips, include=FALSE}

#to infer fishing activities:

df2<-read.table("example_data_fishing.txt",sep=",")

```
