---
title: "Potential code flow"
subtitle: "data-examples/doi_10.5061_dryad.k80bp46__v1.txt"
author: "Einar Hjörleifsson"
date: "`r lubridate::today()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Preambles

```{r all_stuff_needed}
library(sf)
library(lubridate)
library(tidyverse)
source("R/ramb_functions.R")
source("R-dev/einar/ramb_functions2.R")
```

Tend to like short-names (less typing donwstream):
```
vid:  vessel id
time: the ais clock
hid:  harbour id
gid:  gear id
tid:  trip id
```
## Preprocessing

### Get data

```{r data}
hb <- 
  read_rds("data/harbours.rds") %>% 
  # give each harbour a unique id
  mutate(hid = 1:n()) %>% 
  select(hid)
d <- 
  read.table("data-examples/doi_10.5061_dryad.k80bp46__v1.txt", 
             sep = ",") %>% 
  as_tibble() %>% 
  # date is date, time is time
  rename(time = date) %>% 
  mutate(time = ymd_hms(time)) %>% 
  # id and ID are the same
  select(-ID) %>% 
  # not needed but less recoding
  rename(vid = id)
glimpse(d)
nrow0 <- nrow(d)  # sanity check downstream
```

### Ping in harbour and trip assignment

```{r distinct_harbour_and_trips}
d <- 
  d %>% 
  # just to be sure
  arrange(vid, time) %>% 
  # may be of use downstream
  mutate(.rid = 1:n()) %>% 
  select(.rid, everything())
print(paste("rows in: ", nrow0, " rows out: ", nrow(d)))
# time needs to be distinct ----------------------------------------------------
d <-
  d %>% 
  # time has to be unique witin a vessel
  #  may want to investigate why this happens and also which duplicate point to keep
  distinct(vid, time, .keep_all = TRUE) 
print(paste("rows in: ", nrow0, " rows out: ", nrow(d)))
# point in polygon (vessel in harbour) and trips -------------------------------
d <-
  d %>% 
  # get harbour id
  st_as_sf(coords = c("x", "y"),
           crs = '+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0',
           remove = FALSE) %>% 
  st_transform(crs = 4326) %>% 
  mutate(lon = st_coordinates(.)[ , 1],
         lat = st_coordinates(.)[ , 2]) %>% 
  st_join(hb) %>% 
  st_drop_geometry() %>% 
  mutate(tid = rb_define_trips(vid, hid))
print(paste("rows in: ", nrow0, " rows out: ", nrow(d)))
d %>% count(vid, tid)
```

```{r trip_overview}
d %>% 
  count(vid, tid) %>% 
  spread(vid, n) %>% 
  knitr::kable(caption = "Overview of trips by vessels (negative values are harbour 'trips').")
```

So, only 1 trip per vessel. Nothing more to do

### Whacky points

```{r whacky_points}
# whacky points ----------------------------------------------------------------
d <- 
  d %>% 
  group_by(vid, tid) %>% 
  mutate(n = n()) %>% 
  #     step speed (nmi)
  mutate(ss = ifelse(n > 2 & tid > 0,
                     rb_arcdist(lat, lon, lead(lat), lead(lon)) / as.numeric(difftime(lead(time), time, units = "hours")),
                     NA_real_)) %>%
  # vmask may crash if put together with other things inside a mutate
  #  even then, seems to be a bit shaky
  #  slow, even for reasonable amount of ais data [volunteering for code optimization??]
  #  need a miminum of six points for vmask, see help file to get info what functions does
  mutate(wacky = ifelse(n >= 6 & tid > 0,
                        argosfilter::vmask(lat, lon, time, vmax = 13),  # m/s ~ 25 knots
                        NA)) %>% 
  ungroup() %>% 
  select(-n)
print(paste("rows in: ", nrow0, " rows out: ", nrow(d)))
```

```{r}
d %>% count(wacky)
```

So nice and neat data :-)

```{r whacky_peek, fig.height = 9}
# lets just plot the thing in any case
d %>% 
  # just for visuals
  mutate(ss = ifelse(ss > 30, 30, ss),
         wacky = ifelse(wacky == "removed", TRUE, FALSE),
         size = ifelse(wacky, 1, 0.01)) %>% 
  ggplot(aes(time)) +
  theme_bw() +
  #geom_point(aes(y = speed), size = 1, colour = "grey") +
  geom_point(aes(y = ss, colour = as_factor(wacky), size = size),
             alpha = 0.5) +
  scale_size_area(max_size = 2) +
  facet_wrap(~ vid, scales = "free_x", ncol = 3) +
  scale_colour_brewer(palette = "Set1") +
  theme(legend.position = "none") +
  labs(x = NULL, y = "kn")
```

No whacky's, no time gaps so 1 trip per vessel.

## Assigning pings to behaviour

The code presented here is inspired by the supplement to [“Identifying fishing grounds from vessel tracks: model-based inference for small scale fisheries” by Tania Mendo, Sophie Smout, Theoni Photopoulou and Mark James (2019) Royal Society Open Science](https://doi.org/10.1098/rsos.191161). The supplementary document provides a code for "five different methods for identifying hauling versus non-hauling activities in small scale fishing vessels from their movement trajectories."

A [sample dataset](https://doi.org/10.5061/dryad.k80bp46) of movement data collected every 60sec from 5 trips by 5 different small scale fishing vessels using creels is provided and used here (has been included in the {ramb}-package).

The objective here is to provide a more structured code than provided in the supplement, relying fully on tidyverse lingo including usage of the {purrr} map-function family, rather than loops. Those details are though hidden within the function-calls, the aim of the default output those that are needed for vessel track behavioural postprocessing analysis.

### Modeling

```{r}
d.change <- 
  d %>% 
  group_by(vid) %>% 
  mutate(change = rb_event(behaviour)) %>% 
  group_by(vid, change, behaviour) %>% 
  summarise(t1 = min(time),
            t2 = max(time),
            .groups = "drop") %>% 
  # make a continuum, not loose the last point
  group_by(vid) %>% 
  mutate(t2 = lead(t1)) %>% 
  ungroup()
p.base <- 
  d.change %>% 
  ggplot() +
  theme_bw() +
  geom_rect(aes(xmin = t1, xmax = t2, ymin = -Inf, ymax = Inf,
                fill = behaviour),
            show.legend = FALSE) +
  facet_wrap(~ vid, scales = "free") +
  scale_fill_manual(values = c("steaming" = "grey",
                               "hauling" = "pink",
                               "shooting" = "green")) +
  ggnewscale::new_scale_fill()
```


#### Model 1: Gaussian mixture model

... not run

#### Model 2: Trip-based Gaussian mixture model

```{r}
d2 <- 
  d %>% 
  rb_gaussian(vid, lon, lat) %>% 
  mutate(Fishing = ifelse(.speed <= threshold.upper, TRUE, FALSE))
d2 %>%
  count(behaviour, Fishing) %>%
  spread(Fishing, n) %>%
  knitr::kable()
d2 %>%
  ggplot(aes(.speed, fill = Fishing)) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~ vid, scales = "free_y") +
  theme(legend.position = c(0.8, 0.2)) +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "Speed [kt]", y = "Pings")
p.base +
  geom_point(data = d2,
             aes(time, .speed, colour = Fishing),
             size = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Speed [kt]") +
  theme(legend.position = c(0.8, 0.2))
```

#### Model 3: Trip-based Binary Clustering using Gaussian mixture models

```{r}
d3 <- 
  d %>% 
  group_by(vid) %>% 
  rb_gaussian_binary_clustering(vid, time, lon, lat) %>% 
  mutate(Fishing = ifelse(.A %in% 1:2, TRUE, FALSE))
d3 %>%
  count(behaviour, .A) %>%
  spread(.A, n) %>%
  knitr::kable()
d3 %>%
  count(behaviour, Fishing) %>%
  spread(Fishing, n) %>%
  knitr::kable()
d3 %>%
  ggplot(aes(.speed, .turn, colour = factor(.A))) +
  theme_bw() +
  geom_point(size = 1) +
  facet_wrap(~ vid) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Speed [kn]", y = "Turn", colour = "clusters") +
  theme(legend.position = c(0.8, 0.25))
d3 %>%
  ggplot(aes(time, .speed)) +
  theme_bw() +
  geom_point(aes(colour = factor(.A)),
             size = 1) +
  facet_wrap(~ vid, scales = "free") +
  theme(legend.position = c(0.8, 0.2)) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Speed [kt]")
p.base +
  geom_point(data = d3,
             aes(time, .speed, colour = Fishing),
             size = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Speed [kt]") +
  theme(legend.position = c(0.8, 0.2))
d3 %>%
  ggplot(aes(time, .turn)) +
  theme_bw() +
  geom_point(aes(colour = factor(.A)),
             size = 1) +
  facet_wrap(~ vid, scales = "free") +
  theme(legend.position = c(0.8, 0.2)) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Turn")
p.base +
  geom_point(data = d3,
             aes(time, .turn, colour = Fishing),
             size = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Turn") +
  theme(legend.position = c(0.8, 0.2))
d3 %>%
  arrange(desc(.A)) %>%
  ggplot(aes(lon, lat, colour = factor(.A))) +
  theme_bw() +
  geom_point(size = 1) +
  facet_wrap(~ vid, scales = "free") +
  coord_quickmap() +
  labs(x = NULL, y = NULL, colour = "clusters") +
  scale_colour_brewer(palette = "Set1") +
  theme(legend.position = c(0.8, 0.25))  +
  scale_x_continuous(NULL, NULL) +
  scale_y_continuous(NULL, NULL)
d3 %>%
  arrange(desc(.A)) %>%
  ggplot(aes(lon, lat, colour = Fishing)) +
  theme_bw() +
  geom_point(size = 1) +
  facet_wrap(~ vid, scales = "free") +
  coord_quickmap() +
  labs(colour = "Fishing") +
  scale_colour_brewer(palette = "Set1") +
  theme(legend.position = c(0.8, 0.25)) +
  scale_x_continuous(NULL, NULL) +
  scale_y_continuous(NULL, NULL)
```

#### Model 4: Hidden Markov model with speed only

where is the speed here?

```{r}
d4 <- 
  d %>% 
  rb_hidden_markov_step(vid, time, x, y)
d4 %>%
  count(behaviour, .vit) %>%
  spread(.vit, n) %>%
  knitr::kable()
p.base +
  geom_point(data = d4,
             aes(time, .step, colour = factor(.vit)),
             size = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Step [m]") +
  theme(legend.position = c(0.8, 0.2))
d4 %>%
  mutate(vid = paste(vid, behaviour)) %>%
  ggplot(aes(time, .step, colour = factor(.vit))) +
  geom_point() +
  facet_wrap(~ vid, scales = "free", ncol = 3) +
  scale_colour_brewer(palette = "Set1")
# why not on speed??
```

#### Model 5: Hidden Markov Model with speed and turning angle

```{r}
d5 <- 
  d %>% 
  rb_hidden_markov_step_and_turn(vid, time, x, y)
d5 %>%
  count(behaviour, .vit) %>%
  spread(.vit, n) %>%
  knitr::kable()
p.base +
  geom_point(data = d5,
             aes(time, .step, colour = factor(.vit)),
             size = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Time", y = "Step [m]") +
  theme(legend.position = c(0.8, 0.2))
d5 %>% 
  mutate(id = paste(vid, behaviour)) %>%
  ggplot(aes(time, .step, colour = factor(.vit))) +
  geom_point() +
  facet_wrap(~id, scales = "free", ncol = 3) +
  scale_colour_brewer(palette = "Set1")
```
